Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset

### 2025-07-15 code/manuscript.Rmd
updaded review for manuscript and updated readme file

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou ^[Center for the Built Environment, University of California Berkeley, USA], Paul Raftery ^1^, Stefano Schiavon ^1^ ^[Department of Architecture, College of Environmental Design, University of California, Berkeley, USA] ^[Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA], Carlos Duarte ^1^, Gail Brager ^1^
abstract: Conventional measurement and verification (M&V) methods rely on comparing pre- and post-retrofit performance to estimate energy savings. They are often time-consuming and unreliable, especially when non-routine events, as step changes or more gradual changes in how a building operates, occur during the M&V process. Those events are unrelated to the intervention but significantly affect building energy consumption and thus when the analyst applies the conventional M&V method, the results can be confounded. In this study, we demonstrated that switchable interventions, such as most HVAC control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction). Additionally, we found that when non-routine events are present, the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

@@

Measurement and Verification (M&V) is the process of quantifying energy
savings from energy efficiency projects by comparing actual energy
consumption against a baseline, adjusting for factors like weather and
occupancy. This process ensures that improvements in energy performance
are accurately evaluated. In the United States, practitioners often
refer to ASHRAE Guideline 14, the International Performance Measurement
and Verification Protocol (IPMVP), and the Federal Energy Management
Program (FEMP) for standard guidelines [@doe_mv_2008;
@efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level, where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. In conventional practice, the process
begins with baseline measurements taken during the year before
implementing the energy-efficiency retrofit, followed by the same
measurement procedure during the year after the intervention. After
collecting two years of data, an M&V analyst fits an energy prediction
model, using variables such as outdoor temperature and time
[@mathieu_quantifying_2011] to project baseline energy consumption in
the post-retrofit period. The difference between the counter-factual
baseline and the measured intervention period energy consumption
represents the energy savings. One of the key drawbacks of this method
is its reliance on a two-year timeline to quantify savings, during which
baseline measurements can become outdated due to changes in building
performance caused by non-routine events unrelated to the intervention.
Specifically, an M&V analyst should account for two types of baseline
changes: (1) static changes, which are known building operational
changes: such as renovation, equipment addition or removal, and changes
related to tenant turnover; and (2) gradual changes, which often occur
subtly and unknown, such as the incremental adoption of LED lighting or
more efficient plug loads. Identifying and quantifying these changes are
particularly labor-intensive and challenging. Consequently, those
limitations reduce the feasibility of M&V and complicate the
quantification of estimation uncertainty, thus impacting the accuracy
and timeliness of savings assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston_comparison_2019; @duan2013;
@gupta2019; @mie2022; @munro; @raseduzzaman2017]. An early application
of this randomized design to the field of M&V in buildings was first
used in [@raftery2018], which compared the energy performance of two
different supply air temperature reset strategies in a building.
Subsequent improvements to the M&V method include adding a sequential
evaluation framework and defining stopping criteria to end the M&V
period early after reaching a target level of uncertainty, using blocked
randomization by weekday. The full framework is detailed in a previously
published study [@raftery2024] along with a case study application to a
real building, and other example use cases. In summary, this method
provides M&V analysts with a randomized schedule that alternates between
baseline and intervention implementation while ensuring balanced
sampling across days of the week and seasons. For example, given a
10-week M&V period for one intervention, the balanced randomized
schedule would equally sample five Mondays with the baseline strategy in
operation and five Mondays with the intervention strategy in operation.
The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off, which is a small subset of all the interventions performed
in buildings. However, for all applicable use cases, it allows analysts
to detect energy savings sequentially shortly after the test begins. In
addition, once the desired estimate uncertainty target is achieved,
analysts can terminate the M&V study and switch to 100% intervention.
The key advantage of randomization is that if control strategies are
sampled with equal probability, the influence of other confounding
factors such as occupancy change (a common non-routine event) and other
more subtle long term changes in building energy use are likely to be
evenly distributed among measurements, leading to a more accurate and
unbiased assessment of the intervention effect.

@@

To demonstrate and differences between the two M&V methods, the Building
Genome Dataset 2 (BGD2) is used, which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies, acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and realistic operational
measurements between 2016 and 2017 from across North America and Europe,
making it one of the most comprehensive collections of building-related
data available for scientific use. The dataset includes various
commercial building types such as offices, education facilities, public,
and retail buildings, and provides detailed information on their
physical characteristics (e.g. energy ratings, heating types and floor
area) and hourly measurements of chilled and hot water, electricity, gas
usage as well as site outdoor weather conditions. In this study, we
mostly used the metered electricity and outdoor weather measurements for
evaluating the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required. In
addition, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have CV(RMSE) lower than
25% for 12-month data collection. Thus, most research related to M&V for
the whole-building approach focuses on the accuracy of baseline
modeling, exploring model performance from simple regression models to
more complex machine learning techniques. One study reviewed various
models suitable for M&V applications as well as selected input features
[@alrobaie_review_2022] and another study provided a definitive
methodology to apply machine learning models for M&V use cases
[@gallagher_development_2018]. In addition, a few studies investigated
the critical performance metrics to evaluate the developed baseline
models [@granderson_automated_2015; @granderson_development_2014] and
compared a variety of models using those metrics
[@granderson_accuracy_2016]. These studies made significant
contributions by emphasizing the uncertainty associated with the
model-fitting process, a key factor in accurately determining energy
savings. Furthermore, other researchers addressed this issue by
leveraging statistical formulation and inference to improve baseline
energy models [@burkhart_measurement_2014; @heo_calibration_2012;
@walter_uncertainty_2014]. However, a gap still remains in the
literature regarding the rigorous quantification of uncertainties
directly associated with calculated savings, for instance, accounting
for the potential bias that the baseline model might deteriorate (i.e.,
become 'stale') over the extended period of pre- and post-analysis.

@@

Changes in a building that influence its energy usage and are unrelated
to the intervention being studied by the M&V method are commonly
referred to as non-routine events. Current guidelines define non-routine
events as 'static factors' that need to be adjusted after projecting the
baseline in the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a general
approach for consideration. For example, IPMVP requires the facility
owner and the M&V analyst to periodically perform inspections of all
equipment and operations during the reporting period, which is
labor-intensive and error-prone. ASHRAE Guideline 14 recommends
performing engineering calculations or computer software simulations to
adjust the post-retrofit baseline. Additionally, it is relatively rare
to have access to all measurements needed for adjustments, such as
occupancy, and thus the analyst normally assumes those factors remain
unchanged throughout the study. For example, some studies use linear
interpolation to estimate counter-factual baseline for the demand
response program [@beil_round-trip_2015; @keskar_commercial_2020].
Consequently, one study points out that it is inaccurate to assume no
change in the operating conditions during the response period (Huang,
Katipamula, and Lutes 2023). Other studies realized such limitation in
field study, and emphasized the importance of requiring matched groups
to control for exogenous factors beyond weather differences when
comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mostly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. We found a similar concept briefly defined in
Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during multi-year
M&V project. However, currently there is no detailed instruction
provided. If M&V analysts are unaware of such changes and lack an
appropriate adjustment method (e.g., replacing filters before the
intervention begins), the savings could be overestimated as the
decreased energy use is incorrectly attributed to the intervention
rather than the filter replacement. An operational drift has not been
formally defined in the literature or standards, but in this study, we
consider it as a special type of non-routine event, as it also
represents a change in facility operation that occurs gradually over a
longer period of time.

@@

As mentioned, the goal of an M&V project is to determine the effect,
typically energy savings, of some intervention in the building. In this
study, we focus on switchable interventions and an example of such an
intervention is a control strategy that adjusts the chilled water
plant's supply water temperature, which can be commonly found in the
literature [@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. In our case, we
made it even simpler by adjusting the setpoint based on outdoor weather
conditions (see description in Section
\@ref(apply-virtue-intervention)). By conducting such an analysis, we
aim to:

1.  Determine, using a large sample of buildings, how much more
    accurately the randomized M&V method would estimate the savings of
    the proposed intervention compared to the conventional M&V, and how
    much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset. The BDG2 dataset
    contains real-world building energy usage data collected over two
    years, which inherently contains varying degrees of usage changes.
    By virtually introducing intervention effects into the existing
    dataset, we can calculate the ground-truth energy savings, which
    serve as a reference point for method comparison.

3.  Open-source implementation of the proposed randomized M&V method. We
    ensured the reproducibility of the method by making the analysis
    code open source, including randomized schedule generation,
    sequential statistical analysis, energy modeling and normalized
    saving calculation. In addition, we also included code for extended
    use cases of the randomized method, such as changing sampling ratio
    and sampling intervals. Using the available open resources, building
    analysts should be able to seamlessly integrate and apply them in
    their own M&V projects.

@@

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section.

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="70%"}

@@

We extracted the electricity measurements from the BDG2 dataset and
filter out all qualified buildings based on the following criteria:

@@

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage in the second year should be
    less than 25% of that in the first year.

@@

For all qualified buildings, we also create a feature to identify
buildings with very stable electricity usage overall by identifying
those with no statistically significant difference (P-value \> 0.05)
between their two-year electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings. While we did not adjust for weather
differences between the two years when generating this high-level
feature, this result highlights that the assumption that buildings are
typically stable over the timeframe involved in conventional M&V methods
is rarely true. As mentioned in \@ref(non-routine-events-impact), such
variability is largely associated with non-routine events, especially
operational drifts since the assessment, even under ideal situations,
lasts for two years. And according to the statistics shown here, it is a
typical case observed in real buildings applications.

## Apply virtue intervention

Figure \@ref(fig:chwst) shows the algorithm for the proposed control
intervention that resets the chiller supply temperature based on the
outdoor weather conditions, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy, as illustrated in the
figure, adjusts the water supply temperature dynamically, resetting it
from 7°C to 12°C.

```{r chwst, fig.cap = "Proposed intervention strategy: chilled water supply temperature reset based on outdoor temperature.", fig.width=8, fig.height=4}

data.frame(
  OAT = c(10, 15, 25, 30),  
  Baseline = c(6, 6, 6, 6),      
  Intervention = c(12, 12, 7, 7)
  ) %>% 
  pivot_longer(cols = c("Baseline", "Intervention"), names_to = "strategy", values_to = "value") %>% 
  ggplot(aes(x = OAT, y = value, color = strategy)) +
  geom_line(linewidth = 1.2) +
  scale_color_manual(values = ls_colors) +
  scale_y_continuous(expand = c(0, 0), 
                     breaks = c(4, 6, 8, 10, 12),
                     limits = c(5, 12.5),
                     labels = number_format(suffix = " °C")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = c(5, 10, 15, 20, 25, 30), 
                     labels = number_format(suffix = " °C")) +
  geom_vline(xintercept = 10, lty = "dashed", color = "red") +
  annotate(geom = "text", x = 7.5, y = 8, label = "No chiller\noperation\nbelow 10 °C", size = 5) +
  coord_cartesian(xlim = c(5, 30)) +
  labs(x = "Outdoor temperature", 
       y = "Chilled water supply temperature", 
       color = NULL, 
       title = "Chilled water supply temperature reset intervention") +
  theme(panel.grid.major.y = element_line(color = "grey80", linewidth = 0.25),
        legend.direction = "horizontal",
        legend.position = "bottom",
        plot.margin = margin(t = 2, r = 7, b = 2, l = 2, unit = "mm"))

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings: $$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc_savings}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
further consumes around 50% of the HVAC electricity. While this
assumption largely simplifies the diverse energy usage across various
building types, for the scope of this paper, we assume that 25% of the
total building electricity is used by the chilled water plant, $E_{CHW}$
[@us2012commercial]. Typically, the savings from an intervention are not
proportional to the building's hourly electricity usage, which is
generally the challenge for M&V. To address this, we mapped the
resulting electricity savings as a percentage of the plant's normal
operation, calculated as its mean electricity usage over the two-year
period. This percentage is influenced by factors such as outdoor
temperature ($OAT$), intervention supply water temperature
($T_{interv}$), baseline supply water temperature ($T_{base}$) and hour
of the day ($\delta_{occ}$, binary indicator whether during peak hours
from 9 AM to 4 PM).

$$
\text{perc_savings = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below. For simplicity, those parameters were not rigorously calibrated
for each building and were applied uniformly in the filtered dataset.
Interested readers can also change the parameters in our open-source
code to simulate different scenarios.

| Parameter              | Description                                                      | Value |
|----------------|-----------------------------------------|----------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

## Run M&V methods

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post-comparison on a 12-month baseline and 12-month
intervention timeline. The randomized method defines sampling
requirements as:

-   Use a daily sampling interval with the sampling time at midnight
    each day.

-   Block by day of the week with a block period of 12 weeks.

and stopping criteria as:

>>>>

---
title: "Demonstrating the Reliability of Randomized Measurement and Verification for Switchable Control Retrofits Using a Large Open-Source Dataset"
author:
  - Aoyu Zou:
      institute: cbe
      correspondence: true
  - name: Paul Raftery
    institute: cbe
  - name: Stefano Schiavon
    institute: [cbe, arch, cee]
  - name: Carlos Duarte
    institute: cbe
  - name: Gail Brager
    institute: [cbe, arch]
institute:
  - cbe: Center for the Built Environment, University of California Berkeley, USA
  - arch: Department of Architecture, College of Environmental Design, University of California, Berkeley, USA
  - cee: Department of Civil and Environmental Engineering, College of Engineering, University of California, Berkeley, USA
abstract: Conventional measurement and verification (M&V) methods for estimating energy savings rely on comparing pre- and post-retrofit performance. They are often time-consuming and unreliable, especially when non-routine events, such as step changes or more gradual changes in building operation, occur during the M&V process. When those events are unrelated to the retrofit intervention but may significantly affect building energy consumption, then when the analyst applies the conventional M&V method the results can be confounded. In this study, we demonstrated that switchable interventions, such as most HVAC control retrofits, can benefit from a new M&V method that randomly samples whether to implement the baseline or the intervention strategy at a fixed interval (e.g., daily). We tested this novel randomized M&V method on a large public dataset (639 buildings) covering various climate zones and commercial building types, using a virtual chilled water supply temperature reset based on outdoor weather as the intervention. The results show that, compared to the conventional method, the randomized method provides more accurate savings estimations with a median of 74% accuracy improvement and is even faster (typically 36 weeks instead of 104 weeks, ~65% reduction). Additionally, we found that when non-routine events are present (e.g., occupancy pattern change), the randomized method estimates savings that are much closer to the ground-truth savings than the conventional method, demonstrating much improved reliability. We also assessed the impact of normalizing for different weather, starting the M&V at different dates of the year, continuing randomization with a different sampling ratio after satisfying all criteria, and dropping samples affected by carryover effects when switching between strategies. For each scenario, we identified the optimal sampling interval using the large dataset.

output:
  bookdown::word_document2:
    reference_docx: "../paper/template.docx"
    number_sections: false
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'

scholarly-metadata:
  affiliations-location: footnote

editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
biblio-style: apalike

@@

Measurement and Verification (M&V) is the process of quantifying
savings, most commonly energy savings, but also potentially water and
demand savings, from efficiency projects by comparing actual consumption
against an established baseline, while adjusting for varying factors
such as weather and occupancy. This process ensures that improvements in
energy performance are accurately evaluated. In the United States,
practitioners often refer to ASHRAE Guideline 14, the International
Performance Measurement and Verification Protocol (IPMVP), and the
Federal Energy Management Program (FEMP) for standard guidelines
[@doe_mv_2008; @efficiency_valuation_organisation_international_2012;
@ashrae_ashrae_2023]. These guidelines outline standardized methods for
quantifying energy savings, whether through calibrated simulations or
monitored measurements for specific equipment or systems (isolation
methods) or for entire buildings (whole-building methods). In this
study, we will focus on the energy savings quantified at the
whole-building level, where the measurements are obtained from utility
bills or whole-building meters, however the overall conclusions would
also apply to isolation methods. In conventional practice, the process
begins with baseline measurements taken during the year before
implementing the energy-efficiency retrofit, followed by the same
measurement procedure during the year after the intervention. After
collecting two years of data, an M&V analyst fits an energy prediction
model, using variables such as outdoor temperature and time
[@mathieu_quantifying_2011] to project baseline energy consumption in
the post-retrofit time period (but without the retrofit itself). The
difference between this counter-factual baseline and the measured energy
consumption during the actual post-retrofit (intervention) period is the
energy savings. However, other changes unrelated to the implemented
strategies, such as occupancy patterns, can also contribute to the
energy consumption difference. Although building analysts can apply
routine adjustments to account for expected variations in independent
variables commonly included in energy models, such as weather conditions
and time, it is hard to accurately adjust for building performance
change caused by non-routine events. Those events refer to less
predictable and infrequent events that impact energy consumption but are
unrelated to the specific intervention under study. Specifically, when
using the conventional method, an M&V analyst should account for two
additional types of baseline changes beyond routine adjustments: (1)
known static changes to building operation, which are considered as
typical non-routine events: including renovation, equipment addition or
removal, and changes related to tenant turnover; and (2) gradual
changes, which often occur subtly and unknown due to lack of accurate
measurements to quantify, such as the incremental adoption of LED
lighting or more efficient plug loads. Identifying and quantifying these
changes are particularly labor-intensive and challenging. Consequently,
those limitations reduce the feasibility of M&V and complicate the
quantification of estimation uncertainty, thus impacting the accuracy
and timeliness of savings assessments.

@@

To address the limitations of conventional M&V methods and the
challenges posed by building baseline changes, we proposed a M&V method
that adopts the randomized crossover design [@raftery2024], a gold
standard from medical, agricultural and online controlled experimental
(i.e. A/B testing) studies [@blackston_comparison_2019; @duan2013;
@gupta2019; @mie2022; @munro; @raseduzzaman2017]. An early application
of this randomized design to the field of M&V in buildings was first
used in Raftery et al. [@raftery2018], which compared the energy
performance of two different supply air temperature reset strategies in
a building. Through random sampling, the study reached a conclusion of
29% energy cost savings after just 6 months. Subsequent improvements to
the M&V method included adding a sequential evaluation framework and
defining stopping criteria to end the M&V period early after reaching a
target level of uncertainty, using blocked randomization by weekday. The
full framework is described in a previously published study
[@raftery2024] along with a case study application to a real building,
and other example use cases. In summary, this method provides M&V
analysts with a randomized schedule that alternates between baseline and
intervention implementation while ensuring balanced sampling across days
of the week and seasons. For example, given a 10-week M&V period for one
intervention, the balanced randomized schedule would equally sample five
Mondays with the baseline strategy in operation and five Mondays with
the intervention strategy in operation.

The major limitation of this method is that it is only applicable to a
subset of retrofit projects where interventions can be easily switched
on and off (e.g., HVAC control retrofits), which is a small subset of
all the interventions performed in buildings. However, for all
applicable use cases, it allows analysts to detect energy savings
sequentially shortly after the test begins. In addition, once the
desired estimate uncertainty target is achieved, analysts can terminate
the M&V study and switch to 100% intervention. The key advantage of
randomization is that when control strategies are sampled with equal
probability, the influence of confounding factors (e.g., occupancy
change---a common non-routine event---and other more subtle long term
changes in building energy use) are likely to be evenly distributed
among measurements, leading to a more accurate and unbiased assessment
of the intervention effect.

@@

To demonstrate the differences between the two M&V methods, we used the
Building Genome Dataset 2 (BGD2), which is an extensive open-access
dataset designed to advance research and development in building energy
efficiency and control strategies, acting as a test-bed for modeling,
simulation, and algorithm development [@miller_building_2020]. BGD2
contains over 1000 buildings' metadata and real operational measurements
between 2016 and 2017 from across North America and Europe, making it
one of the most comprehensive collections of building-related data
available for scientific use. The dataset includes various commercial
building types such as offices, education facilities, public, and retail
buildings, and provides detailed information on their physical
characteristics (e.g. energy ratings, heating types and floor area) and
hourly measurements of chilled and hot water, electricity, gas usage as
well as site outdoor weather conditions. In this study, we mostly used
the metered electricity and outdoor weather measurements for evaluating
the M&V methods.

@@

ASHRAE Guideline 14 and the IPMVP provide options for whole-building M&V
use cases. For code compliance using the prescriptive option, 12 months
of baseline and 12 months of post-retrofit measurements are required.
Moreover, the expected savings should be larger than 10% and the
baseline energy model fitting accuracy should have a Coefficient of
Variation of the Root Mean Square Error, CV(RMSE), lower than 25% for
12-month data collection. Thus, most research related to M&V for the
whole-building approach focuses on the accuracy of baseline modeling,
exploring model performance from simple regression models to more
complex machine learning techniques. Alrobaie and Krarti
[@alrobaie_review_2022] reviewed various models suitable for M&V
applications as well as selected input features. Gallagher et al.
[@gallagher_development_2018] provided a well-thought-out methodology to
apply machine learning models for M&V use cases. Researchers at LBNL
investigated the critical performance metrics to evaluate the developed
baseline models [@granderson_automated_2015;
@granderson_development_2014] and compared a variety of models using
those metrics [@granderson_accuracy_2016]. These studies emphasized the
uncertainty associated with the model-fitting process, a key factor in
accurately determining energy savings. Furthermore, other researchers
addressed this issue by leveraging statistical formulation and inference
to improve baseline energy models [@burkhart_measurement_2014;
@heo_calibration_2012; @walter_uncertainty_2014]. However, a gap still
remains in the literature regarding how to rigorously quantify the
uncertainties directly associated with calculated savings. For instance,
this includes accounting for the potential bias that the baseline model
might deteriorate over time, as well as the model no longer accurately
representing the building energy behavior due to changes in operation,
occupancy, equipment performance, etc.

@@

As mentioned earlier, non-routine events are changes in a building that
influence its energy usage but are unrelated to the intervention being
studied by the M&V. Current guidelines define non-routine events as
'static factors' that need to be adjusted after projecting the baseline
into the post-retrofit period [@ashrae_ashrae_2023;
@efficiency_valuation_organisation_international_2012]. A common
non-routine event in energy-saving M&V projects is a change in occupancy
or a significant shift in occupant behavior, equipment run time, and
operating conditions (e.g. set points, lighting, and ventilation
levels). However, current standards or guidelines only provide a
reminder that non-routine events should be reported and properly
adjusted. For example, IPMVP requires the facility owner and the M&V
analyst to periodically perform inspections of all equipment and
operations during the reporting period, which is labor-intensive and
error-prone. ASHRAE Guideline 14 recommends performing engineering
calculations or computer software simulations to adjust the
post-retrofit baseline. Additionally, it is relatively rare to have
access to all measurements needed for adjustments, such as occupancy,
and thus the analyst normally assumes those factors remain unchanged
throughout the study. For example, some studies use linear interpolation
to estimate counter-factual baseline for the demand response program
[@beil_round-trip_2015; @keskar_commercial_2020]. Consequently, it is
inaccurate to assume no change in the operating conditions during the
response period [@huang_experimental_2023]. This is a limitation in a
field study, and it is important to require matched groups to control
for exogenous factors beyond weather (which is usually taken into
account) when comparing between baseline and intervention
[@demand_side_analytics_population_2022; @huang_experimental_2023].

Overall, those non-routine events mainly refer to clearly observable
changes, but there are also more gradual and subtle changes in
buildings, which are hardly noticed. We define those gradual changes as
'operational drift' in this study. One likely encountered operational
drift situation is the filter clogging in air handling units due to
particle accumulation. This causes supply fans to gradually consume more
energy to maintain the required duct static pressure [@feng_newly_2019;
@zhai_full-scale_2017]. We found a similar concept briefly defined in
Guideline 14 as the 'load creep' [@ashrae_ashrae_2023] during a
multi-year M&V project. However, currently there are no detailed
instructions provided on how to account for them. If M&V analysts are
unaware of such changes and lack an appropriate adjustment method (e.g.,
replacing filters before the intervention begins), the savings could be
overestimated as the decreased energy use is incorrectly attributed to
the intervention rather than the filter replacement. An operational
drift has not been formally defined in the literature or standards; in
this study, we consider it as a special type of non-routine event, as it
also represents a change in facility operation that occurs gradually
over a longer period of time.

@@

One goal of an M&V project is to determine the effect, typically energy
savings, of an intervention in the building. In this study, we focus on
switchable interventions. An example of such an intervention is a
control strategy that strategically adjusts the chilled water plant's
supply water temperature with proven ability to save energy
[@duarte_field_2023; @lee_simulationoptimization_2012;
@qiu_chilled_2022; @jin2007; @taylor_optimizing_2012]. We used a simple
version of the control strategy in which the setpoint is adjusted based
on outdoor temperature (see description in Section
\@ref(apply-virtual-intervention)). By conducting such an analysis, we
aim to:

1.  Determine how much more accurately the randomized M&V method would
    estimate the savings of the intervention compared to the
    conventional M&V, and how much more quickly it would reach a result.

2.  Determine how much more robust the randomized method is compared to
    the conventional method when static non-routine events and
    operational drift are present in the dataset.

3.  Make the implementation of the proposed randomized M&V method
    open-source.

The significance of this study is that the BDG2 dataset contains
real-world building energy usage data collected over two years, which
inherently contains varying degrees of usage changes. In addition, by
virtually introducing intervention effects into the existing dataset, we
can calculate the ground-truth energy savings, which serve as a
reference point for method comparison. Furthermore, we ensured the
reproducibility of the method by making the analysis code open source,
including randomized schedule generation, sequential statistical
analysis, energy modeling and normalized saving calculation. This also
covers the extended discussion below on sensitivity analysis and other
use cases of the randomized method, such as changing sampling ratio and
sampling intervals. Using the available open resources, building
analysts should be able to seamlessly integrate and apply them in their
own M&V projects.

@@

We outlined the methodology of the study in Figure \@ref(fig:flowchart)
and extended several key components in this section. As a summary, we
utilized the historical measurements from a dataset that contains a
large number of real commercial buildings as the pre-retrofit baseline
and simulated an intervention strategy. Then, we applied both
conventional and randomized M&V methods to the modified dataset to
estimate intervention savings, and compare which one produces results
closer to the simulated intervention effect.

```{r flowchart, fig.cap = "Flow chart showing the methodology for comparing the estimated savings of randomized M&V with the conventional M&V", out.width="60%"}

@@

We extracted the electricity measurements from the BDG2 dataset and
filtered out all qualified buildings based on the following criteria:

@@

3.  $\frac{abs(E_{2017} - E_{2016})}{E_{2016}}< 25\%$: any increase or
    decrease in building electricity usage ($E$) in the second year
    ($E_{2017}$) should be less than 25% of that in the first year
    ($E_{2016}$).

@@

As a result, we kept a total of 639 buildings from 10 different
locations for the subsequent analysis, and we included a more detailed
summary of the building features in the supplementary material. For all
qualified buildings, we also create a feature to identify buildings with
very stable electricity usage overall by identifying those with no
statistically significant difference (P-value \> 0.05) between their
pre- and post-retrofit electricity usage. Applying this filter to real
buildings, only 66 buildings are labeled in this 'stable' subset out of
a total of more than 600 buildings, corresponding to 11%. While we did
not adjust for weather differences between the two years when generating
this high-level feature, this result highlights that the assumption that
buildings are typically stable over the timeframe involved in
conventional M&V methods is rarely true; in 89% of the cases this
assumption is not valid. As mentioned in Section
\@ref(non-routine-events-impact), such variability is largely associated
with non-routine events---especially operational drifts---since the
assessment, even under ideal situations, lasts for two years. According
to the statistics shown here, it is a typical case observed in real
buildings applications.

## Apply **virtual** intervention

The proposed control intervention resets the chiller supply temperature
based on the outdoor temperature, which can be commonly found in the
literature [@lee_chilled_2022; @congradac_recognition_2012]. For both
strategies, we assume that the chiller operates when the outdoor
temperature exceeds 10°C. The baseline strategy, representing the
existing measurements from the dataset, operates with a constant water
supply temperature. The intervention strategy adjusts the water supply
temperature dynamically, resetting it from 7°C to 12°C. We included a
visual summary of the algorithm in the supplementary material.

We used a simple relationship to map the chilled water supply
temperature reset to the electrical energy savings:$$
\text{savings} = mean(E_{CHW}) \times 25\% \times \text{perc}
$$

We assume on average, HVAC systems account for approximately 50% of a
building's total electricity consumption, and the chilled water plant
$E_{CHW}$ further consumes around 50% of the HVAC electricity, or 25% of
the total building electricity [@us2012commercial]. We recognize that
this assumption, used for the scope of this analysis, largely simplifies
the diverse energy usage across various building types. Typically, the
savings from an intervention are not proportional to the building's
hourly electricity usage, which is generally a challenge for M&V. To
address this, we mapped the resulting electricity savings as a
percentage of the plant's normal operation, calculated as its mean
electricity usage over the two-year period. This percentage is
influenced by factors such as outdoor temperature ($OAT$), intervention
supply water temperature ($T_{interv}$), baseline supply water
temperature ($T_{base}$) and hour of the day ($\delta_{occ}$, binary
indicator whether during peak hours from 9 AM to 4 PM).

$$
\text{perc = }
\begin{cases}
0, & \text{ if } OAT < 10°C \\
\left[ \beta_{\text{temp}} \times (\text{T}_{\text{interv}} - \text{T}_{\text{base}}) \right] \times 
\left[ \beta_{\text{occ}} \times \delta_{occ} + \beta_{\text{unocc}} \times (1 - \delta_{occ})\right], & \text{ otherwise}
\end{cases}
$$

Parameters and their pre-defined values are summarized in the table
below, they were obtained from a trial supply water setpoint reset
experiment on a campus office building [@zou_quantifying_2025]. For
simplicity, those parameters were not rigorously calibrated for each
building and were applied uniformly in the filtered dataset. Interested
readers can also change the parameters in our open-source code to
simulate different scenarios.

| Parameter              | Description                                                      | Value |
|------------------|------------------------------------|------------------|
| $\beta_{temp}$         | \% savings from setting $T_{interv}$ 1 °C higher than $T_{base}$ | 0.08  |
| $\beta_{occ}$          | \% savings adjustment during occupied hours                      | 1.2   |
| $\beta_{\text{unocc}}$ | \% savings adjustment during unoccupied hours                    | 0.8   |

@@

## M&V methods and results analysis

We described in detail the workflow of both conventional and randomized
M&V methods in the previous study [@raftery2024]. The conventional
method is a pre- and post-comparison on a 12-month baseline and 12-month
intervention timeline. The randomized method defines sampling
requirements as (a) use a daily sampling interval with the sampling time
at midnight each day and (b) block by day of the week with a block
period of 12 weeks.
